---
title: "Practice for data.table"
subtitle: "with 7 questions"  
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(data.table)
input <- if (file.exists("flights14.csv")) {
  "flights14.csv"
} else {
  "https://raw.githubusercontent.com/Rdatatable/data.table/master/vignettes/flights14.csv"
}
# fast-and-friendly file reader 
flights <- fread(input)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
xaringanExtra::use_tile_view()
xaringanExtra::use_panelset()
xaringanExtra::use_editable()
library(xaringanthemer)
style_solarized_dark( base_font_size = "20px",
  text_font_size = "1.5rem",
  code_font_size = "1rem",
  header_h1_font_size = "2.5rem",
  header_h2_font_size = "2rem",
  header_font_google = google_font("Roboto"),
  text_font_google = google_font("Open Sans"),
  code_font_google = google_font("IBM Plex Mono"),
  header_color = "#FFFFFF")
```

# Before starting...

1. You can write your own answer in each question  

2. Don't forget to import library data.table
```{r}
library(data.table)
```
3. We will use flights data in **nycflights13** library and remove NA

```{r}
library(nycflights13)
data("flights")
flights <- as.data.table(flights) |> na.exclude()
```


---

class: inverse, center, middle

# Basic Syntax (3 questions)
----

---
# Question 1
----

Get all the flights in the May 15th

.panelset[
.panel[.panel-name[Your Answer]
.can-edit[  
]]
.panel[.panel-name[Our Answer]
```{r}
ans <- flights[month == 5 & day == 15]
```

]]
---

# Question 2
----

Sort flights first by column 'month' in ascending order, and then by 'day' in decending order  
.panelset[
.panel[.panel-name[Your Answer]
.can-edit[  
]]
.panel[.panel-name[Our Answer]
```{r}
ans <- flights[order(month, -day)]
```

]]

---

# Question 3
----
Select both arr_delay and dep_delay columns

.panelset[
.panel[.panel-name[Your Answer]
.can-edit[  
]]
.panel[.panel-name[Our Answer]
```{r}
ans <- flights[, .(arr_delay, dep_delay)]
head(ans)
```
]]

---

# Question 4
----

How many trips have had total delay(arr_delay + dep_delay) > 0?

.panelset[
.panel[.panel-name[Your Answer]
.can-edit[  
]]
.panel[.panel-name[Our Answer]
```{r}
ans <- flights[, sum( (arr_delay + dep_delay) > 0 )]
ans

```

]]
---

# Question 5
----

Calculate the average arrival and departure delay for all flights in the may 15th

.panelset[
.panel[.panel-name[Your Answer]
.can-edit[  
]]
.panel[.panel-name[Our Answer]
```{r}
ans <- flights[month == 5 & day == 15,
               .(m_arr = mean(arr_delay), m_dep = mean(dep_delay))]
ans

```
]]
---

# Question 6
----

How can get the number of trips corresponding to each origin airport?

.panelset[
.panel[.panel-name[Your Answer]
.can-edit[  
]]
.panel[.panel-name[Our Answer]
```{r}
ans <- flights[, .(.N), by = .(origin)]
ans

```
]]

---

# Question 7

How can we get 
